Covariate analysis of ROSMAP Reprocessed RNASeq data with TMM

### 2016-09-04 code/Rmd/ROSMAP_RNASeq_Reprocessed_TMM.Rmd
Added markdown to process ROSMAP with voom TMM normalisation

---
title: "Covariate analysis of ROSMAP Reprocessed RNASeq"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./ROSMAP_RNASeq_Reprocessed.Rmd",
                                 parentId ="syn5570291",
                                 entityName = 'ROSMAP RNASeq Reprocessed Counts (Based on cogdx)')
```

@@

thisFileName <- 'ROSMAP_RNASeq_Reprocessed.Rmd'

@@

Remove genes that have less than 1 cpm counts in at least 50% of samples.
```{r cpmnormalisation, cache=TRUE}
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT,MIN_GENE_CPM=1, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
```
### Normalisation (with NULL)
Initial normalisation usign voom (with NULL design)
```{r initial.voom.normalisation}
# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(PROCESSED_COUNTS$filteredExprMatrix, design=NULL, plot=T)
```
### Outlier analysis
Clustering of voom adjusted data (with NULL design)
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale.=T, center = T)

@@

  # Estimate correlation of random effects
  VOOM.ADJUSTED.GENE_EXPRESSION = voom(PROCESSED_COUNTS$filteredExprMatrix, design=DM1$design, plot=F)

@@

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`
### Sanity check
```{r residualSigCovars.manual, fig.width=20, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "cogdx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("cogdx", colnames(DM1$design), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

@@

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC)) + geom_point() + xlim(c(-1,1))
p = p + geom_hline(yintercept = -log10(0.05), color = 'red') 
p = p + geom_vline(xintercept = log2(1.2), color = 'red') 
p = p + geom_vline(xintercept = -log2(1.2), color = 'red') 
p = p + facet_grid(.~Comparison, scales = 'fixed')
p
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ROSMAP RNASeq Reprocessed Counts (Based on cogdx)", parentId = parentId)
CODE <- synStore(CODE)

>>>>

---
title: "Covariate analysis of ROSMAP Reprocessed RNASeq data with TMM"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./ROSMAP_RNASeq_Reprocessed_TMM.Rmd",
                                 parentId ="syn5570291",
                                 entityName = 'ROSMAP RNASeq Reprocessed Counts TMM (Based on cogdx)')
```

@@

thisFileName <- 'ROSMAP_RNASeq_Reprocessed_TMM.Rmd'

@@

Remove genes that have less than 1 cpm counts in at least 50% of samples.
```{r cpmnormalisation, cache=TRUE}
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT,MIN_GENE_CPM=1, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
```
### Normalisation (with NULL)
Initial normalisation usign voom (with NULL design)
```{r initial.voom.normalisation}
# TMM normalisation
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix)

# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=NULL, plot=T)
```
Coexpression of genes 
```{r coexp1, cache=FALSE}
cr = cor(t(VOOM.GENE_EXPRESSION$E))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
### Outlier analysis
Clustering of voom adjusted data (with NULL design)
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale.=T, center = T)

@@

  # Estimate correlation of random effects
  VOOM.ADJUSTED.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=DM1$design, plot=F)

@@

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`
### Sanity check
```{r residualSigCovars.manual, fig.width=20, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "cogdx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("cogdx", colnames(DM1$design), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC)) + geom_point() + xlim(c(-1,1))
p = p + geom_hline(yintercept = -log10(0.05), color = 'red') 
p = p + geom_vline(xintercept = log2(1.2), color = 'red') 
p = p + geom_vline(xintercept = -log2(1.2), color = 'red') 
p = p + facet_grid(.~Comparison, scales = 'fixed')
p
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ROSMAP RNASeq Reprocessed Counts TMM (Based on cogdx)", parentId = parentId)
CODE <- synStore(CODE)
